---
title: "MEC with blocking"
author: "Adam Struzik"
output: 
    html_vignette:
        df_print: kable
        toc: true
        number_sections: true
        fig_width: 6
        fig_height: 4
vignette: >
  %\VignetteIndexEntry{MEC with blocking}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Setup

Load required packages.

```{r setup, message = FALSE}
library(automatedRecLin)
library(data.table)

options("text2vec.mc.cores" = 1L)
```

## Data

We use the full example Census and Customer Information System (CIS) datasets
from [McLeod et al. (2011)](https://wayback.archive-it.org/12090/20231221144450/https://cros-legacy.ec.europa.eu/content/job-training_en).
The goal is to link records from CIS to records from Census.

```{r data}
data("census", package = "automatedRecLin")
data("cis", package = "automatedRecLin")
setDT(census)
setDT(cis)

NROW(cis)
NROW(census)
```

The `person_id` variable identifies the correct linkage.
We use this information only to evaluate the result.

```{r true-matches}
cis[is.na(cis)] <- ""
census[is.na(census)] <- ""

cis[, pername1 := gsub("-", "", pername1)]
census[, pername1 := gsub("-", "", pername1)]

true_matches <- merge(
  x = cis[, .(a = .I, person_id)],
  y = census[, .(b = .I, person_id)],
  by = "person_id"
)[, .(a, b)]

NROW(true_matches)
```

## MEC with blocking

We compare forename and surname using the Jaro-Winkler distance. These two
comparison variables are modeled with the continuous parametric MEC method.
Sex and date-of-birth variables use the default binary method. Address fields
are used only to construct blocks.

```{r model-specification}
variables <- c(
  "pername1", "pername2", "sex",
  "dob_day", "dob_mon", "dob_year"
)

comparators <- list(
  "pername1" = jarowinkler_complement(),
  "pername2" = jarowinkler_complement()
)

methods <- list(
  "pername1" = "continuous_parametric",
  "pername2" = "continuous_parametric"
)

blocking_variables <- c(variables, "enumcap", "enumpc")
```

Run blocked MEC. The model is trained on sampled blocks that contain at least
the requested number of pairs and a lower bound on nonmatches.

```{r mec-blocking}
set.seed(1)

result <- mec_blocking(
  A = cis,
  B = census,
  variables = variables,
  comparators = comparators,
  methods = methods,
  blocking_variables = blocking_variables,
  blocking_sep = "",
  controls_blocking = list(seed = 1, n_threads = 1),
  min_training_pairs = 1000,
  min_training_nonmatches = 1000,
  block_sampling_seed = 1,
  nonmatch_sample_size = 100000,
  nonmatch_sampling_seed = 1,
  true_matches = true_matches
)

result
```

## What blocking buys us

The full Cartesian product contains `r format(NROW(cis) * NROW(census), big.mark = ",")`
record pairs. Blocking reduces this to `r format(result$blocking_eval[["blocked_pairs"]], big.mark = ",")`
candidate pairs, while retaining `r sprintf("%.2f%%", 100 * result$blocking_eval[["blocking_recall"]])`
of known links. The final linkage set contains `r format(NROW(result$M_est), big.mark = ",")`
predicted matches.

```{r results, echo = FALSE}
data.table(
  step = c("Training", "Blocking", "Linkage"),
  result = c(
    paste0(
      result$training_rule, " on ",
      format(NROW(result$training_blocks), big.mark = ","),
      " blocks"
    ),
    paste0(
      format(result$blocking_eval[["preserved_matches"]], big.mark = ","),
      " of ",
      format(result$blocking_eval[["true_matches"]], big.mark = ","),
      " known links retained"
    ),
    paste0(
      "FLR = ", sprintf("%.2f%%", 100 * result$eval_metrics[["FLR"]]),
      "; MMR = ", sprintf("%.2f%%", 100 * result$eval_metrics[["MMR"]])
    )
  )
)
```
